---
title: "gtsummary in quarto"
format:
  html:
    theme: default
execute: 
  echo: true
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(gtsummary)
nlsy_cols <- c("glasses", "eyesight", "sleep_wkdy", "sleep_wknd",
               "id", "nsibs", "samp", "race_eth", "sex", "region",
               "income", "res_1980", "res_2002", "age_bir")
nlsy <- read_csv(here::here("data", "raw", "nlsy.csv"),
                 na = c("-1", "-2", "-3", "-4", "-5", "-998"),
                 skip = 1, col_names = nlsy_cols) |>
  mutate(region_cat = factor(region, labels = c("Northeast", "North Central", "South", "West")),
         sex_cat = factor(sex, labels = c("Male", "Female")),
         race_eth_cat = factor(race_eth, labels = c("Hispanic", "Black", "Non-Black, Non-Hispanic")),
         eyesight_cat = factor(eyesight, labels = c("Excellent", "Very good", "Good", "Fair", "Poor")),
         glasses_cat = factor(glasses, labels = c("No", "Yes")))
```

# Customization of `tbl_summary()`

We’re using some data from the **National Longitudinal Survey of Youth 1979**, a cohort of American young adults aged 14-22 at enrollment in 1979. They continue to be followed to this day, and there is a wealth of publicly available data online. I’ve downloaded the answers to a survey question about whether respondents wear glasses, a scale about their eyesight with glasses, their (NLSY-assigned 😒) race/ethnicity, their sex (presumably at birth, but unclear), their family’s income in 1979, and their age at the birth of their first child.

```{r}
#| label: tbl-nlsy-summary
#| tbl-cap: "Summary of NLSY participant characteristics by sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

@tbl-nlsy-summary presents a summary of demographic and health characteristics of NLSY participants by sex.

```{r}
#| label: tbl-nlsy-characteristics-labeled
#| tbl-cap: "Characteristics of NLSY participants by sex with descriptive variable labels"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir),
  label = list(
    race_eth_cat ~ "Race/ethnicity",
    region_cat ~ "Region",
    eyesight_cat ~ "Eyesight",
    glasses ~ "Wears glasses",
    age_bir ~ "Age at first birth"
  ),
  missing_text = "Missing")
```

@tbl-nlsy-characteristics-labeled summarizes the demographic and health characteristics of NLSY participants by sex, with descriptive variable labels and missing values displayed.

##table practice

```{r}
#| label: tbl-nlsy-characteristics
#| tbl-cap: "Characteristics of NLSY participants by sex"
tbl_summary(
  nlsy,
  by = sex_cat,
  include = c(sex_cat, race_eth_cat, region_cat,
              eyesight_cat, glasses, age_bir))
```

@tbl-nlsy-characteristics summarizes the demographic and health characteristics of NLSY participants by sex.

```{r}
#| label: tbl-sleep-summary
#| tbl-cap: "Characteristics of NLSY participants"

tbl_summary(
  nlsy,
  include = c(sex_cat, race_eth_cat, sleep_wkdy, sleep_wknd),
  label = list(
    sex_cat ~ "Sex",
    race_eth_cat ~ "Race/ethnicity",
    sleep_wkdy ~ "Weekday sleep",
    sleep_wknd ~ "Weekend sleep"
  )
)

```

@tbl-sleep-summary shows the descriptive statistics for weekend sleep among NLSY participants.

```{r}
median_sleep <- median(nlsy$sleep_wknd, na.rm = TRUE)
```

The median number of hours of sleep on weekends was `r median_sleep` hours.

The median number of hours of sleep on weekends was `r median(nlsy$sleep_wknd, na.rm = TRUE)` hours.

```{r}
sleep_table <- tbl_summary(
  nlsy,
  include = sleep_wknd
)
```

The median number of hours of sleep on weekends was `r inline_text(sleep_table, variable = sleep_wknd, pattern = "{median}")` hours.

```{r}
sleep_table <- tbl_summary(
  nlsy,
  include = c(sleep_wkdy, sleep_wknd),
  statistic = all_continuous() ~ "{median} ({p25}, {p75})"
)
```

```{r}
##sleep_table$table_body
inline_text(sleep_table, variable = "sleep_wknd",
						 column = "stat_0", pattern = "{median}, IQR = [{p25}, {p75}]")
```

In the sample the average sleep on weekends was `r inline_text(sleep_table, variable = "sleep_wknd",column = "stat_0", pattern = "{median}, IQR = [{p25}, {p75}]")`

```{r}

```

